---
title: {{ default "Comparison of cluster resolutions with `clustree`" (getenv "TITLE") }}
weight:

categories:

tags:

authors:
  - christopher-barrington

type: docs
toc: true
draft: false

date: `r format(Sys.time(), "%Y-%m-%d")`
created: {{ now.Format "2006-01-02" }}

always_allow_html: true
output:
  html_document:
    keep_md: yes
    df_print: paged
---

<!-- initial set up chunks are invisible -->

<!-- read the standard knitr setup script -->
```{r knitr_setup}
#| code=readLines(system.file('chunks/knitr.r', package='project.notes.tools')),
#| echo=FALSE
```

<!-- make subdirectories to write into/any bash work -->
```{bash setup_wd}
#| code=readLines(system.file('chunks/working_directory.sh', package='project.notes.tools')),
#| opts.label='batch_chunk'
```

<!-- load any packages for the workbook -->
```{r load_workbook_packages}
#| opts.label='batch_chunk'

library(clustree)
library(Seurat)
```

<!-- read the standard setup script, loading well-used libraries -->
```{r configure_r}
#| code=readLines(system.file('chunks/environment.r', package='project.notes.tools')),
#| opts.label='batch_chunk'
```

<!-- set a ggplot theme to use as a base in the workbook -->
```{r define_global_ggplot_theme}
#| opts.label='batch_chunk'

(theme_bw() +
 theme(text=element_text(size=10, colour='black'),
       panel.grid.minor=element_blank(),
       strip.background=element_blank())) %>%
theme_set()
```

<!-- any final things to set for this workbook -->
```{r final_setup}
#| opts.label='batch_chunk'

doMC::registerDoMC(32)
```

<!-- define slugs for other reports here for ease and reuse -->
```{r write_slugs}
#| child=system.file('children/write_slugs.Rmd', package='project.notes.tools'),
#| slugs=list()
```

[CRICK]: https://www.crick.ac.uk

[clustree home]: https://github.com/lazappi/clustree
[clustree github plotting issue]: https://github.com/lazappi/clustree/issues/63

<!-- end of initial set up chunks -->

Cell clusters have been defined in the previous workbook(s) at multiple resolutions. Here, we compare the cells in each cluster at each resolution to decide which resolution to use. The analysis compares the cell IDs to determine which clusters become split by increasing resolution or which clusters remain stable with higher resolutions.

<!--more-->

## Run `clustree` 

We use [`clustree`][clustree home] to analyse the clusters in each Seurat object. The result is a `ggplot` object that can be manipulated as any other `ggplot` object.

Each `rds` Seurat object is read directly into `clustree` in the following chunk; there is no need to save the object in memory.

The following function takes a path to a Seurat object RDS file and runs `clustree` on the `meta.data` slot.

{{< callout "note" >}}
The number of resolutions considered is reduced in these plots for interpretation. Resolutions above 3.5 are omitted from the analysis.
{{< /callout >}}

```{r function_to_get_clustree_plots}
run_clustree <- function(rds) {
  seurat <- readRDS(rds)

  seurat %>%
    slot(name='meta.data') %>%
    select(-ends_with('_snn_res.3.5'):-seurat_clusters) -> obj

  seurat %>%
    DefaultAssay() %>%
    str_c('_snn_res.') -> prefix

  rm('seurat')
  gc()

  tryCatch(expr={clustree(x=obj, prefix=prefix, layout='tree', show_axis=TRUE)},
           error=function(x) {clustree(x=obj, prefix=prefix, layout='tree', show_axis=FALSE)}) +
    labs(y='Cluster resolution',
         size='Cells (n)',
         edge_colour='Cells (n)',
         edge_alpha='Proportion cells',
         colour='Cluster resolution') +
    guides(colour='none') -> result

    result
}
```

The plots show each cluster (`x`) and resolution (`colour` and `y`) and the number of cells in the cluster (`size`). Between cluster resolutions the number of cells contributing to the higher resolution cluster is shown by the colour and transparency of the arrow. An opaque yellow indicates a high number and proportion of cells in the lower-resolution cluster form the cells in the higher-resolution cluster. 

{{< callout "notice" >}}
Using these plots a cluster resolution can be found at which the clusters become stable and further increasing resolution has little effect on cluster content. This may be the resolution at which cell types become most-resolved.
{{< /callout >}}

<!--  -->

## QC-filtered datasets

```{r get_clustree_plots_for_qc_filtered}
#| dependson='function_to_get_clustree_plots',
#| echo=TRUE

data(project_sample_sheet)

file.path('../../seurat', 'load-quantified-matrices', 'scratch/clustering/out/seurat') %>%
  list.files(pattern='*rds', full.names=TRUE) %>%
  set_names() %>%
  set_names(basename) %>%
  set_names(str_remove, pattern='.rds$') %>%
  enframe(name='project', value='path') %>%
  inner_join(x=project_sample_sheet, by='project') %>%
  dlply(~project_label, .parallel=TRUE, function(x) run_clustree(rds=x$path)) -> plots_list
```

{{< tabs "show_clustree_plots_for_qc_filtered" >}}
```{r show_clustree_plots_for_qc_filtered}
#| cache=FALSE,
#| dependson='get_clustree_plots_for_qc_filtered',
#| echo=FALSE,
#| results='asis'

for(i in names(plots_list)) {
  sprintf('{{< tab "%s" >}}', i) %>% cat()
  plots_list[[i]] %>% resize_and_show(width=4, height=9)
  cat('{{< /tab >}}')
}
```
{{< /tabs >}}

<!--  -->

## Subsets from QC-filtered datasets

```{r get_clustree_plots_for_qc_filtered_subsets}
#| dependson='function_to_get_clustree_plots',
#| echo=TRUE

data(project_sample_sheet)

file.path('../../seurat', 'subset-qc-filtered-datasets', 'scratch/clustering/out') %>%
  data.frame(path=., rds=list.files(path=., recursive=TRUE, full.names=FALSE)) %>%
  mutate(path=file.path(path, rds)) %>%
  separate(col=rds, into=c('cell_set', 'outtype', 'project'), sep='/') %>%
  filter(outtype=='seurat') %>%
  mutate(project=str_remove(project, '.rds$')) %>%
  inner_join(x=project_sample_sheet, by='project') %>%
  dlply(~project_label+cell_set, .parallel=TRUE, function(x) run_clustree(rds=x$path)) -> plots_list
```

{{< tabs "show_clustree_plots_for_qc_filtered_subsets" >}}
```{r show_clustree_plots_for_qc_filtered_subsets}
#| dependson='get_clustree_plots_for_qc_filtered_subsets',
#| echo=FALSE,
#| results='asis'

plots_list %>%
  attr(which='split_labels') %>%
  dlply(~project_label, pluck, 'cell_set') -> tabs_list

for(project in names(tabs_list)) {
  sprintf(fmt='{{< tab "%s" >}}\n', project) %>% cat()
  sprintf(fmt='{{< tabs "show_clustree_plots_for_qc_filtered_subsets-%s" >}}\n', project) %>% cat()

  for(cell_subset in tabs_list[[project]]) {
    sprintf('{{< tab "%s" >}}', cell_subset) %>% cat()
    plots_list %>% pluck(str_c(project, cell_subset, sep='.')) %>% resize_and_show(width=4, height=9)
    cat('{{< /tab >}}\n')
  }

  cat('{{< /tabs >}}  \n')
  cat('{{< /tab >}}')
}
```
{{< /tabs >}}

<!-- include a references section; DOIs for software used in the workbook or to generate presented results, a named vector of strings -->
```{r write_bibliography}
#| bibliography_doi=c(clustree='10.1093/gigascience/giy083'),
#| child=system.file('children/write_bibliography.Rmd', package='project.notes.tools')
```

<!-- include a session information section -->
```{r show_session_information}
#| child=system.file('children/session_information.Rmd', package='project.notes.tools'),
#| opts.label='batch_chunk',
#| saved_objects <- c()
```
